Oriented Gradient Histogram applied to P300 Detection

### 2017-11-29 article.tex
Mas cambios en las fórmulas

On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot.   Around the keypoint, a local image patch is constructed by dividing it in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and $ \mathbf{kp} = (x_{kp},y_{kp})$ is the patch center.

A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the $16$ blocks and creating a histogram of gradients.  This technique is based on Lowe's SIFT~\citep{Lowe2004} method, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.   In order to calculate the histogram, the interval $[0-360]$ of possible angles is divided in $8$ bins, each one at $45$ degrees.
Then, the angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $. For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.

\begin{equation}
 h(\theta,i,j) = 3s \sum_{\mathbf{x}} w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{x})|
\label{eq:histogram}
\end{equation}

\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al. in~\citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained. In each step, the histogram is computed by multiplying by the length of the gradient vector, in order to take account of it.

Fig.~\ref{fig:sampledescriptor} shows an example of a patch and a scheme of the histogram computation. Fig.~\ref{patch} is a plot of the signal and the patch centered in the keypoint. In Fig.~\ref{histogrambines} the orientations on each patch are illustrtaed which encompass the corresponding $\mathbf{kp}$-descriptor of $128$ coordinates. The first two bloks orientations are shown.

\begin{figure}[H]
\centering
\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.5\linewidth]{DiagramaConkeypoint.png}\label{patch}}
\subfigure[Orientations on each patch.]{\includegraphics[width=.5\linewidth]{histogramsbines.png}\label{histogrambines}}
%\includegraphics[width=9cm]{gradientszoom.png}
\caption{ Example of a patch and a scheme of the orientations histogram computation.}
\label{fig:sampledescriptor}
\end{figure}

@@

While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the image that were used to derive those descriptors.

The NBNN algorithm tackles this problem by comparing each image against a whole classification class which is characterized by the set of descriptors that are closest to each one of the descriptors obtained for the query image.  The algorithm is based on

\begin{equation}
\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
\label{eq:classification}
\end{equation}

\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.

>>>>

On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered. A local image patch is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center.

A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the $16$ blocks and creating a histogram of gradients.  This technique is based on Lowe's SIFT~\citep{Lowe2004} method, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.   In order to calculate the histogram, the interval $[0-360]$ of possible angles is divided in $8$ bins, each one at $45$ degrees.
Then, the angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $. For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.

\begin{equation}
 h(\theta,i,j) = 3s \sum_{\mathbf{x} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{x})|
\label{eq:histogram}
\end{equation}

\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al. in~\citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained. It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.

Fig.~\ref{fig:sampledescriptor} shows an example of a patch and a scheme of the histogram computation. Fig.~\ref{patch} is a plot of the signal and the patch centered in the keypoint. In Fig.~\ref{histogrambines} the possible orientations on each patch are illustrated. They form the corresponding $\mathbf{kp}$-descriptor of $128$ coordinates. The first two blocks orientations are shown.

\begin{figure}[H]
\centering
\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{DiagramaConkeypoint.png}\label{patch}}
\subfigure[Orientations on each patch.]{\includegraphics[width=.49\linewidth]{histogramsbines.png}\label{histogrambines}}
%\includegraphics[width=9cm]{gradientszoom.png}
\caption{ Example of a patch and a scheme of the orientations histogram computation.}
\label{fig:sampledescriptor}
\end{figure}

@@

While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the whole image. %that were used to derive those descriptors.

The NBNN algorithm tackles this problem by comparing each image against a whole classification class. It is characterized by the set of descriptors closest to each one of the descriptors obtained for the query image.  The algorithm  find the predicted class $\hat{C}$ of a query image by means of the following minimization:

\begin{equation}
\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
\label{eq:classification}
\end{equation}
where $C$ is the a class, $ d_i $ is a descriptor that belongs to the query image and $ NN_C(d_i) $ is the corresponding near neighbor descriptor for the class C.
%\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
